---
title: "Practical Machine Learning Final Project"
author: "Yuchen Ding"
date: "1/3/2021"
output: html_document
---

## Executive Summary
  The goal of this project is to predict the manner (5 different ways) of 6 participants have performed, by using data collected from accelerometers on the belt, forearm, arm, and dumbell. Three different models (decision tree, random forest, and GBM) were used to train the data and the one with best predicting accuracy was selected to predict the 20 test cases available in the test data.

## Background
  Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Preparing the Environment
```{r cache = T}
setwd("E:/study/coursera/Data Science Specialization/Course 8 Practical Machine Learning")
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(corrplot)
library(randomForest)
library(gbm)
```

## Loading the Data
```{r cache = T}
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
validUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainRaw <- read.csv(url(trainUrl))
validRaw <- read.csv(url(validUrl))
dim(trainRaw)
dim(validRaw)
```

## Data Preprocessing

#### Split the Data for Training
```{r}
set.seed(7) 
inTrain <- createDataPartition(trainRaw$classe, p = 0.7, list = F)
trainingRaw <- trainRaw[inTrain, ]
testingRaw <- trainRaw[-inTrain, ]
dim(trainingRaw)
dim(testingRaw)
```

#### Removing Variables with lots of NAs (> 90 %)
```{r}
manyNAs <- sapply(trainingRaw, function(x) mean(is.na(x))) > 0.9
training <- trainingRaw[, manyNAs == F]
testing <- testingRaw[, manyNAs == F]
dim(training)
dim(testing)
```

#### Removing Variables with Little Impact on the Outcome
```{r}
training <- training[, -c(1:6)]
testing <- testing[, -c(1:6)]
dim(training)
dim(testing)
```

#### Removing Variables with Near Zero Variance
```{r}
nzVar <- nearZeroVar(training)
training <- training[, -nzVar]
testing <- testing[, -nzVar]
dim(training)
dim(testing)
```

## Correlation Analysis
```{r echo = T}
corMatrix <- cor(training[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "upper", tl.cex = 0.3)
```
As most of the variables are less-correlated, all 53 variables will be used for building the ML models.

## Building the ML models
### Model 1: Decision Trees
```{r echo = T}
set.seed(7)
DT <- rpart(classe ~ ., data = training, method = "class")
fancyRpartPlot(DT)
```
```{r}
DTPred <- predict(DT, testing, type = "class")
DTcm <- confusionMatrix(DTPred, factor(testing$classe))
DTcm
```

### Model 2: Random Forests
```{r echo = T}
set.seed(7)
ctrlRF <- trainControl(method = "cv", number = 3, verboseIter = F)
RF <- train(classe ~ ., data = training, method = "rf", trControl = ctrlRF)
```
```{r}
RFPred <- predict(RF, testing, type = "raw")
RFcm <- confusionMatrix(RFPred, factor(testing$classe))
RFcm
```

### Model 3: Generalized Boosted Machine'
```{r echo = T}
set.seed(7)
ctrlGBM <- trainControl(method = "repeatedcv", number = 3, repeats = 1)
GBM <- train(classe ~ ., data = training, method = "gbm", trControl = ctrlGBM, verbose = F)
```
```{r}
GBMPred <- predict(GBM, testing)
GBMcm <- confusionMatrix(GBMPred, factor(testing$classe))
GBMcm
```

From the above three models, we found that random forests yields the highest accuracy, and we will use it to predict the unkonwn data.

## Applying Random Forest Model to Predict the Test Data
```{r echo = T}
results <- predict(RF, newdata = validRaw)
results
```